---
title: ""
author: "Teresa Ober"
date: "July 8, 2021"
output: 
  github_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      error = FALSE, 
                      message = FALSE,
                      warning = FALSE,
                      plain.ascii = FALSE,
                      fig.fullwidth = TRUE, fig.width = 6.5, fig.asp = .8)
```

```{r, include = FALSE}
library(dplyr); library(tidyverse) # data manipulation
library(psych) # quick descriptive/summary statistics
library(metafor)
library(robumeta)
```

# Getting Started  
Before the workshop, please consider taking a moment to complete the following steps:
1) Download the latest version of R statistical environment from the website here: https://cran.cmm.msu.ru/ .  
2) Though not required, you will probably find it helpful to also download RStudio: https://www.rstudio.com/products/rstudio/download/ . For our purposes, the free "RStudio Desktop" version will work just fine. 
- If you're not familiar with RStudio at all, [Chapter 1 of this online textbook](https://rafalab.github.io/dsbook/getting-started.html) provides a gentle introduction.  
3) During the workshop, we will be using the following packages, so if you are able to install R, please also consider installing them ahead of time. Note that you should only need to install the package once (unless you later re-install or update R). From then on, you can call the package into your library by using the "library()" or "require()" functions.   
-- [*tidyverse*](): Data manipulation  
-- [*dplyr*](): Data manipulation  
-- [*psych*](): Creating quick descriptive summaries of data  
-- [*metafor*](): Meta-analysis   
-- [*robumeta*](): Robust variance estimation meta-analysis  
-- [*ggplot2*](): Plotting  

Most of these packages have packages dependencies (i.e., other packages they use), so you may notice other packages getting installed, as well.  

\newpage
# Resources  
## Preferred Reporting Items for Systematic Reviews and Meta-Analyses (PRISMA)  
- Moher, D., Liberati, A., Tetzlaff, J., Altman, D. G., & PRISMA Group. (2009). Preferred reporting items for systematic reviews and meta-analyses: the PRISMA statement. *PLoS Medicine*, *6*(7), e1000097. https://doi.org/10.1371/journal.pmed.1000097  
- Moher, D., Shamseer, L., Clarke, M., Ghersi, D., Liberati, A., Petticrew, M., ... & Stewart, L. A. (2015). Preferred reporting items for systematic review and meta-analysis protocols (PRISMA-P) 2015 statement. *Systematic Reviews*, *4*(1), 1-9. https://doi.org/10.1186/2046-4053-4-1  

## Pre-registry for Systematic Reviews  
- PROSPERO (For pre-registration of health-related systematic reviews): https://www.crd.york.ac.uk/prospero/  
- Research Registry ($): https://www.researchregistry.com/  

## R Packages for Meta-analysis
- Metafor: https://cran.r-project.org/web/packages/metafor/index.html  
- Robumeta: https://cran.r-project.org/web/packages/robumeta/index.html  

## General Resources for R and RStudio
- https://rafalab.github.io/dsbook


```{r}
# the theory and rationale of meta-analysis
# identifying and developing research questions appropriate for meta-analysis
# creating and protocol for the systematic review and drafting a pre-registration
# accessing and organizing citations
# PRISMA procedure for screening articles (and tools for collaborative article screening)
# procedures for data extraction and establishing inter-coder reliability
# other resources for evaluating the quality of the systematic review
```









