---
title: ""
author: "Teresa Ober"
date: "July 8, 2021"
output: 
  github_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      error = FALSE, 
                      message = FALSE,
                      warning = FALSE,
                      plain.ascii = FALSE,
                      fig.fullwidth = TRUE, fig.width = 6.5, fig.asp = .8)
```

```{r, include = FALSE}
library(dplyr); library(tidyverse) # data manipulation
library(psych) # quick descriptive/summary statistics
library(metafor)
library(robumeta)
```

# Getting Started  
Before the workshop, please consider taking a moment to complete the following steps:  
1) Download the latest version of R statistical environment from the website here: https://cran.cmm.msu.ru/ .    
2) Though not required, you will probably find it helpful to also download RStudio: https://www.rstudio.com/products/rstudio/download/ . For our purposes, the free "RStudio Desktop" version will work just fine.   
- If you're not familiar with RStudio at all, [Chapter 1 of this online textbook](https://rafalab.github.io/dsbook/getting-started.html) provides a gentle introduction.    
3) During the workshop, we will be using the following packages, so if you are able to install R, please also consider installing them ahead of time. Note that you should only need to install the package once (unless you later re-install or update R). From then on, you can call the package into your library by using the "library()" or "require()" functions.     
-- [*tidyverse*](https://www.tidyverse.org/packages/): Data manipulation  
-- [*dplyr*](https://dplyr.tidyverse.org/): Data manipulation  
-- [*psych*](https://cran.r-project.org/web/packages/psych/index.html): Creating quick descriptive summaries of data  
-- [*metafor*](https://www.metafor-project.org/doku.php): Meta-analysis   
-- [*robumeta*](https://cran.r-project.org/web/packages/robumeta/index.html): Robust variance estimation meta-analysis  
-- [*ggplot2*](https://ggplot2.tidyverse.org/): Plotting  

Most of these packages have packages dependencies (i.e., other packages they use), so you may notice other packages getting installed, as well. 

\newpage  
# Overview of Workshop Topics  

## Part 1: Introduction to Systematic Review  

### Aims:  
- Receive overview of systematic review processes  
- Discuss the scope and the formulation of a review question  
- Consider the purpose and format of a protocol  
- Gain familiarity with  systematic searching/screening strategies  
- Learn data extraction, quality assessment, and synthesis  

### Organization:  
1) What is a systematic review?  
2) Best practices for conducting systematic reviews  
3) Question Formulation  
4) Creating a Protocol  
5) Searching for Literature  
6) Screening Literature   
7) Data Extraction  
8) Risk of Bias and Quality Assessment  
9) Synthesis  

### Slides:  
<style>
.responsive-wrap iframe{ max-width: 100%;}
</style>
<div class="responsive-wrap">
<!-- this is the embed code provided by Google -->
  <iframe src="https://docs.google.com/presentation/d/e/2PACX-1vRZs4CArb-zKJG2vS0nGIbILIpJZyIvCx3UkrTq8E6hIWiqVDIwh7u9O7DAAqg7og/embed?start=false&loop=false&delayms=3000" frameborder="0" width="640" height="389" allowfullscreen="true" mozallowfullscreen="true" webkitallowfullscreen="true"></iframe>
<!-- Google embed ends -->
</div>

## Part 2: Introduction to Meta-analysis

### Aims:  
- Receive overview of steps involved in a meta-analysis  
- Discuss effect size estimates and challenges to extraction  
- Gain an orientation to fixed- and random-effects models  
- Experience data extraction and meta-analysis  
- Outline the reporting of meta-analysis findings  


### Organization:  
1) What is a meta-analysis?  
2) Effect size estimates  
3) Fixed- and Random-Effects  
4) Resources for Conducting Meta-analysis  
5) Reporting Results  


### Slides:  
<style>
.responsive-wrap iframe{ max-width: 100%;}
</style>
<div class="responsive-wrap">
<!-- this is the embed code provided by Google -->
  <iframe src="https://docs.google.com/presentation/d/e/2PACX-1vS_9LZvC7JpMsN1k12ERa7cMujLnuY99QVxiEiGX7nZyYVsRyIQ1PgUFbq2Hza9lQ/embed?start=false&loop=false&delayms=3000" frameborder="0" width="640" height="389" allowfullscreen="true" mozallowfullscreen="true" webkitallowfullscreen="true"></iframe>
<!-- Google embed ends -->
</div>

### Activity Materials:
- *Meta-analysis R Script*: 
[Example_Meta_Analysis_Script.R](https://drive.google.com/file/d/18LXOHctVDZMHR-iKHvfexfrjxhClrwNO/)  
- *Data file*: [ober_2020_ef_decoding.csv](https://drive.google.com/file/d/1YykZF5iP0U6003c9UJeSsK-9K4h0-i8F)  



\newpage
# Resources  

## General Resources for R and RStudio
- Irizarry, R. A. (2019). *Introduction to data science: Data analysis and prediction algorithms with R.* https://rafalab.github.io/dsbook  

## General Resources for Conducting Meta-analysis in R  
- Harrer, M., Cuijpers, P., Furukawa, T.A., & Ebert, D.D. (2021). *Doing Meta-Analysis with R: A Hands-On Guide.* Boca Raton, FL and London: Chapmann & Hall/CRC Press. ISBN 978-0-367-61007-4. https://bookdown.org/MathiasHarrer/Doing_Meta_Analysis_in_R/  


## Additional Documentation of R Packages for Meta-analysis
Polanin, J. R., Hennessy, E. A., & Tanner-Smith, E. E. (2017). A review of meta-analysis packages in R. *Journal of Educational and Behavioral Statistics*, *42*(2), 206-242. https://doi.org/10.3102/1076998616674315  

### Metafor
The metafor Package: Tips and Notes. https://metafor-project.org/doku.php/tips  

### Robumeta
Tanner-Smith, E. E., Tipton, E., & Polanin, J. R. (2016). Handling complex meta-analytic data structures using robust variance estimates: A tutorial in R. *Journal of Developmental and Life-Course Criminology*, *2*(1), 85-112. https://doi.org/10.1002/jrsm.1091  

Fisher, Z., & Tipton, E. (2015). robumeta: An R-package for robust variance estimation in meta-analysis. *arXiv preprint*. https://arxiv.org/abs/1503.02220   

## Pre-registration of Systematic Reviews  
- [PROSPERO](https://www.crd.york.ac.uk/prospero/): For pre-registration of health-related systematic reviews.    
- [Research Registry](https://www.researchregistry.com/): Altnerative to PROSPERO that does not restrict systematic reviews to those related to human health. However, also costs money to submit a pre-registration.    
- Many other alternatives, as well, that are not specific to systematic reviews, including the [Open Science Framework](https://osf.io/prereg/), [AsPredicted.Org](https://aspredicted.org/), etc.

## Preferred Reporting Items for Systematic Reviews and Meta-Analyses (PRISMA)  
Moher, D., Liberati, A., Tetzlaff, J., Altman, D. G., & PRISMA Group. (2009). Preferred reporting items for systematic reviews and meta-analyses: the PRISMA statement. *PLoS Medicine*, *6*(7), e1000097. https://doi.org/10.1371/journal.pmed.1000097  

Moher, D., Shamseer, L., Clarke, M., Ghersi, D., Liberati, A., Petticrew, M., ... & Stewart, L. A. (2015). Preferred reporting items for systematic review and meta-analysis protocols (PRISMA-P) 2015 statement. *Systematic Reviews*, *4*(1), 1-9. https://doi.org/10.1186/2046-4053-4-1  




```{r}
# the theory and rationale of meta-analysis
# identifying and developing research questions appropriate for meta-analysis
# creating and protocol for the systematic review and drafting a pre-registration
# accessing and organizing citations
# PRISMA procedure for screening articles (and tools for collaborative article screening)
# procedures for data extraction and establishing inter-coder reliability
# other resources for evaluating the quality of the systematic review

#For the effect size calculation and moderator/meta-regression analysis, I can plan to give some demos that involve the use of two R packages (metafor, robumeta) for analyzing the data. I should be able to pull some toy data to use for the latter portion to get some hands-on experience running the analyses. If it would be of interest, I'd like to focus on two types of meta-analysis: one which focuses on correlations as effect size estimates, and the other which focuses on group mean differences as effect size estimates.


```









